---
title: "Prepare Data site7"
author: "Natalie"
date: "2024-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
```

#Import all Data

Import all rasters from site 7
```{r load rasters}
site_7_1 <- rast(here("Data", "site7", "7_1_rast.tif"))
names(site_7_1) <- c("red", "green", "blue")
site_7_2 <- rast(here("Data", "site7", "7_2_rast.tif"))
names(site_7_2) <- c("red", "green", "blue")
site_7_3 <- rast(here("Data", "site7", "7_3_rast.tif"))
names(site_7_3) <- c("red", "green", "blue")
site_7_4 <- rast(here("Data", "site7", "7_4_rast.tif"))
names(site_7_4) <- c("red", "green", "blue")
site_7_5 <- rast(here("Data", "site7", "7_5_rast.tif"))
names(site_7_5) <- c("red", "green", "blue")
```

Import all polygons from site 7. Each sub-site contains 5 polygons of each class. 
```{r load polygons}
poly_7_1 <- vect(here("Data", "site7", "7_1_poly.shp"))
poly_7_2 <- vect(here("Data", "site7", "7_2_poly.shp"))
poly_7_3 <- vect(here("Data", "site7", "7_3_poly.shp"))
poly_7_4 <- vect(here("Data", "site7", "7_4_poly.shp"))
poly_7_5 <- vect(here("Data", "site7", "7_5_poly.shp"))
```

#Prepare data

##Generate random points
Create 200 random points within each polygon and save each one as a shapefile
```{r shapefile list}
#Create a list containing all the shapefiles
polygons <- list(
  poly_7_1 = "7_1_poly.shp",
  poly_7_2 = "7_2_poly.shp",
  poly_7_3 = "7_3_poly.shp",
  poly_7_4 = "7_4_poly.shp",
  poly_7_5 = "7_5_poly.shp"
)
```

```{r random point loop}
#Loop through each shapefile, generate 200 random points within each polygon, and save the points as a shapefile. 
for(name in names(polygons)) {
  poly <- vect(here("Data", "site7", polygons[[name]])) #get sub-site polygons
  all_randomPts <- vect() #create empty spatVector to store all random points
  #loop through each polygon in each shapefile
  for(i in 1:nrow(poly)) {
    single_poly <- poly[i, ] #get each individual polygon
    randomPts <- spatSample(single_poly, 200, method = "random") #generate 200 random points in each polygon
    all_randomPts <- rbind(all_randomPts, randomPts) #add random points to the empty spatVector
  }
  subsite_number <- sub("poly_7_", "", name)  # Extract the subsite number
  points_filename <- paste0("7_", subsite_number, "_points.shp") #create filename for the new points dataset
  writeVector(all_randomPts, here("Data", "site7", points_filename), filetype = "ESRI Shapefile")
  print(paste("Finished generating points for site", name))
}
```
##Extract rgb values of the raster at the random points generated
Sample rgb values of the rasters at each sub-site and save the results as a geopackage
```{r raster list}
rasters <- list(
  site_7_1 = site_7_1,
  site_7_2 = site_7_2,
  site_7_3 = site_7_3,
  site_7_4 = site_7_4,
  site_7_5 = site_7_5
)
```

```{r points list}
points <- list(
  pts_7_1 = vect(here("Data", "site7", "7_1_points.shp")),
  pts_7_2 = vect(here("Data", "site7", "7_2_points.shp")),
  pts_7_3 = vect(here("Data", "site7", "7_3_points.shp")),
  pts_7_4 = vect(here("Data", "site7", "7_4_points.shp")),
  pts_7_5 = vect(here("Data", "site7", "7_5_points.shp"))
)
```

```{r extraction loop}
#Loop through each raster and the corresponding points, and extract rgb values
for(raster_name in names(rasters)) {
  subsite_identifier <- sub("site_", "", raster_name)
  points_name <- paste0("pts_", subsite_identifier)
  if (points_name %in% names(points)) {
    raster <- rasters[[raster_name]]
    point_set <- points[[points_name]]
    rgb_values <- extract(raster, point_set)
    point_data <- cbind(point_set, rgb_values)
    gpkg_filename <- paste0(subsite_identifier, "_pts_rgb.gpkg")
    writeVector(point_data, here("Data", "site7", gpkg_filename), filetype = "GPKG")
    print(paste("Finished extracting and saving RGB values for site", subsite_identifier))
  } else {
    print(paste("No matching points file found for raster", raster_name))
  }
}
```


