---
title: "Train Set 1"
author: "Natalie"
date: "2024-08-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
library(dplyr)
```
# Prepare training data
Import all data from subsites 1 to be used for training the model
```{r}
pts_1_1 <- read_sf(here("Data", "site1", "1_1_pts_rgb.gpkg"), stringsAsFactors = T)
pts_2_1 <- read_sf(here("Data", "site2", "2_1_pts_rgb.gpkg"), stringsAsFactors = T)
pts_3_1 <- read_sf(here("Data", "site3", "3_1_pts_rgb.gpkg"), stringsAsFactors = T)
pts_4_1 <- read_sf(here("Data", "site4", "4_1_pts_rgb.gpkg"), stringsAsFactors = T)
pts_5_1 <- read_sf(here("Data", "site5", "5_1_pts_rgb.gpkg"), stringsAsFactors = T)
pts_6_1 <- read_sf(here("Data", "site6", "6_1_pts_rgb.gpkg"), stringsAsFactors = T)
pts_7_1 <- read_sf(here("Data", "site7", "7_1_pts_rgb.gpkg"), stringsAsFactors = T)
```

Select only the columns we want, which are className, red, green, blue and geom. 
```{r}
#Create a list containing all the data
trainSet1_list <- list(pts_1_1, pts_2_1, pts_3_1, pts_4_1, pts_5_1, pts_6_1, pts_7_1)

#Select the columns that we want
select_columns <- function(gpkg) {
  gpkg %>% select(className, red, green, blue, geom)
}

#Ensure all datasets only contain the columns we want
trainSet1_list <- lapply(trainSet1_list, select_columns) 

#Merge all datasets to create the full training dataset
trainSet1 <- do.call(bind_rows, trainSet1_list)
trainSet1
```
# Create model

Create a spatial mlr3 task
```{r}
trainSet1_tsk <- as_task_classif_st(trainSet1, target = "className")
```

Partition the data into training and testing sets 
```{r}
set.seed(100)
trainSet1_split <- partition(trainSet1_tsk, ratio = 0.7)
```

Create random forest learner
```{r}
trainSet1_lrn <- lrn("classif.ranger")
```

Train the learner
```{r}
trainSet1_train <- trainSet1_lrn$train(trainSet1_tsk, trainSet1_split$train)
```

Test the learner's accuracy
```{r}
trainSet1_test <- trainSet1_lrn$predict(trainSet1_tsk, trainSet1_split$test)
trainSet1_test$score(msr("classif.acc"))
```

# Obtain classification accuracy of other subsites in site 1

Import site 1 points with known classes
```{r}
pts_1_2 <- vect(here("Data", "site1", "1_2_points.shp"))
pts_1_3 <- vect(here("Data", "site1", "1_3_points.shp"))
pts_1_4 <- vect(here("Data", "site1", "1_4_points.shp"))
pts_1_5 <- vect(here("Data", "site1", "1_5_points.shp"))
```


Import site 1 rasters for prediction
```{r}
rast_1_2 <- rast(here("Data", "site1", "1_2_rast.tif"))
rast_1_3 <- rast(here("Data", "site1", "1_3_rast.tif"))
rast_1_4 <- rast(here("Data", "site1", "1_4_rast.tif"))
rast_1_5 <- rast(here("Data", "site1", "1_5_rast.tif"))
```

Rename raster layers
```{r}
subsite_ids <- 2:5
for (subsite_id in subsite_ids) {
  raster <- get(paste0("rast_1_", subsite_id))
  names(raster) <- c("red", "green", "blue")
  assign(paste0("rast_1_", subsite_id), raster)
  print(paste("Renamed layers for site 1_", subsite_id))
}
```
Create vector to store accuracy results
```{r}
site1_acc <- numeric(length(subsite_ids))
names(site1_acc) <- paste0("site_1_", subsite_ids)
```

Create a loop that classifies each raster site, extracts the classnames at certain points, and compares them to the known classnames of those points to obtain classification accuracy. 
```{r}
for (i in seq_along(subsite_ids)) {
  subsite_id <- subsite_ids[i]
  raster_to_classify_1 <- get(paste0("rast_1_", subsite_id))
  pts_shapefile_1 <- get(paste0("pts_1_", subsite_id))
  classified_raster_1 <- predict_spatial(raster_to_classify_1, trainSet1_lrn)
  classified_file_path_1 <- here("Data", "site1", paste0("1_", subsite_id, "_classified.tif"))
  writeRaster(classified_raster_1, classified_file_path_1, overwrite=T)
  extracted_vals_1 <- extract(classified_raster_1, pts_shapefile_1)
  raster_className <- extracted_vals_1[,2]
  shapefile_className <- pts_shapefile_1$className
  site1_acc_vect <- raster_className == shapefile_className
  class_acc_1 <- sum(site1_acc_vect) / length(site1_acc_vect)
  site1_acc[i] <- class_acc_1
  print(paste("Classification Accuracy for site1_", subsite_id, ":", class_acc_1))
}
```

```{r}
print(site1_acc)
```

#Obtain classification accuracy of other subsites in site 2
Import site 2 points with known classes
```{r}
pts_2_2 <- vect(here("Data", "site2", "2_2_points.shp"))
pts_2_3 <- vect(here("Data", "site2", "2_3_points.shp"))
pts_2_4 <- vect(here("Data", "site2", "2_4_points.shp"))
pts_2_5 <- vect(here("Data", "site2", "2_5_points.shp"))
```


Import site 2 rasters for prediction
```{r}
rast_2_2 <- rast(here("Data", "site2", "2_2_rast.tif"))
rast_2_3 <- rast(here("Data", "site2", "2_3_rast.tif"))
rast_2_4 <- rast(here("Data", "site2", "2_4_rast.tif"))
rast_2_5 <- rast(here("Data", "site2", "2_5_rast.tif"))
```

Rename raster layers
```{r}
subsite_ids <- 2:5
for (subsite_id in subsite_ids) {
  raster <- get(paste0("rast_2_", subsite_id))
  names(raster) <- c("red", "green", "blue")
  assign(paste0("rast_2_", subsite_id), raster)
  print(paste("Renamed layers for site 2_", subsite_id))
}
```
Create vector to store accuracy results
```{r}
site2_acc <- numeric(length(subsite_ids))
names(site2_acc) <- paste0("site_2_", subsite_ids)
```

Create a loop that classifies each raster site, extracts the classnames at certain points, and compares them to the known classnames of those points to obtain classification accuracy. 
```{r}
for (i in seq_along(subsite_ids)) {
  subsite_id <- subsite_ids[i]
  raster_to_classify_2 <- get(paste0("rast_2_", subsite_id))
  pts_shapefile_2 <- get(paste0("pts_2_", subsite_id))
  classified_raster_2 <- predict_spatial(raster_to_classify_2, trainSet1_lrn)
  classified_file_path_2 <- here("Data", "site2", paste0("2_", subsite_id, "_classified.tif"))
  writeRaster(classified_raster_2, classified_file_path_2, overwrite=T)
  extracted_vals_2 <- extract(classified_raster_2, pts_shapefile_2)
  raster_className <- extracted_vals_2[,2]
  shapefile_className <- pts_shapefile_2$className
  site2_acc_vect <- raster_className == shapefile_className
  class_acc_2 <- sum(site2_acc_vect) / length(site2_acc_vect)
  site2_acc[i] <- class_acc_2
  print(paste("Classification Accuracy for site2_", subsite_id, ":", class_acc_2))
}
print(site2_acc)
```

#Obtain classification accuracy of other subsites in site 3
Import site 3 points with known classes
```{r}
pts_3_2 <- vect(here("Data", "site3", "3_2_points.shp"))
pts_3_3 <- vect(here("Data", "site3", "3_3_points.shp"))
pts_3_4 <- vect(here("Data", "site3", "3_4_points.shp"))
pts_3_5 <- vect(here("Data", "site3", "3_5_points.shp"))
```


Import site 3 rasters for prediction
```{r}
rast_3_2 <- rast(here("Data", "site3", "3_2_rast.tif"))
rast_3_3 <- rast(here("Data", "site3", "3_3_rast.tif"))
rast_3_4 <- rast(here("Data", "site3", "3_4_rast.tif"))
rast_3_5 <- rast(here("Data", "site3", "3_5_rast.tif"))
```

Rename raster layers
```{r}
subsite_ids <- 2:5
for (subsite_id in subsite_ids) {
  raster <- get(paste0("rast_3_", subsite_id))
  names(raster) <- c("red", "green", "blue")
  assign(paste0("rast_3_", subsite_id), raster)
  print(paste("Renamed layers for site 3_", subsite_id))
}
```
Create vector to store accuracy results
```{r}
site3_acc <- numeric(length(subsite_ids))
names(site3_acc) <- paste0("site_3_", subsite_ids)
```

Create a loop that classifies each raster site, extracts the classnames at certain points, and compares them to the known classnames of those points to obtain classification accuracy. 
```{r}
for (i in seq_along(subsite_ids)) {
  subsite_id <- subsite_ids[i]
  raster_to_classify_3 <- get(paste0("rast_3_", subsite_id))
  pts_shapefile_3 <- get(paste0("pts_3_", subsite_id))
  classified_raster_3 <- predict_spatial(raster_to_classify_3, trainSet1_lrn)
  classified_file_path_3 <- here("Data", "site3", paste0("3_", subsite_id, "_classified.tif"))
  writeRaster(classified_raster_3, classified_file_path_3, overwrite=T)
  extracted_vals_3 <- extract(classified_raster_3, pts_shapefile_3)
  print(sum(is.na(extracted_vals_3)))
  raster_className <- extracted_vals_3[,2]
  shapefile_className <- pts_shapefile_3$className
  site3_acc_vect <- raster_className == shapefile_className
  class_acc_3 <- sum(site3_acc_vect) / length(site3_acc_vect)
  site3_acc[i] <- class_acc_3
  print(paste("Classification Accuracy for site3_", subsite_id, ":", class_acc_3))
}
print(site3_acc)
```

#Obtain classification accuracy of other subsites in site 4
Import site 2 points with known classes
```{r}
pts_4_2 <- vect(here("Data", "site4", "4_2_points.shp"))
pts_4_3 <- vect(here("Data", "site4", "4_3_points.shp"))
pts_4_4 <- vect(here("Data", "site4", "4_4_points.shp"))
pts_4_5 <- vect(here("Data", "site4", "4_5_points.shp"))
```


Import site 2 rasters for prediction
```{r}
rast_4_2 <- rast(here("Data", "site4", "4_2_rast.tif"))
rast_4_3 <- rast(here("Data", "site4", "4_3_rast.tif"))
rast_4_4 <- rast(here("Data", "site4", "4_4_rast.tif"))
rast_4_5 <- rast(here("Data", "site4", "4_5_rast.tif"))
```

Rename raster layers
```{r}
subsite_ids <- 2:5
for (subsite_id in subsite_ids) {
  raster <- get(paste0("rast_4_", subsite_id))
  names(raster) <- c("red", "green", "blue")
  assign(paste0("rast_4_", subsite_id), raster)
  print(paste("Renamed layers for site 4_", subsite_id))
}
```
Create vector to store accuracy results
```{r}
site4_acc <- numeric(length(subsite_ids))
names(site4_acc) <- paste0("site_4_", subsite_ids)
```

Create a loop that classifies each raster site, extracts the classnames at certain points, and compares them to the known classnames of those points to obtain classification accuracy. 
```{r}
for (i in seq_along(subsite_ids)) {
  subsite_id <- subsite_ids[i]
  raster_to_classify_4 <- get(paste0("rast_4_", subsite_id))
  pts_shapefile_4 <- get(paste0("pts_4_", subsite_id))
  classified_raster_4 <- predict_spatial(raster_to_classify_4, trainSet1_lrn)
  classified_file_path_4 <- here("Data", "site4", paste0("4_", subsite_id, "_classified.tif"))
  writeRaster(classified_raster_4, classified_file_path_4, overwrite=T)
  extracted_vals_4 <- extract(classified_raster_4, pts_shapefile_4)
  raster_className <- extracted_vals_4[,2]
  shapefile_className <- pts_shapefile_4$className
  site4_acc_vect <- raster_className == shapefile_className
  class_acc_4 <- sum(site4_acc_vect) / length(site4_acc_vect)
  site4_acc[i] <- class_acc_4
  print(paste("Classification Accuracy for site3_", subsite_id, ":", class_acc_4))
}
print(site4_acc)
```

#Obtain classification accuracy of other subsites in site 5
Import site 2 points with known classes
```{r}
pts_5_2 <- vect(here("Data", "site5", "5_2_points.shp"))
pts_5_3 <- vect(here("Data", "site5", "5_3_points.shp"))
pts_5_4 <- vect(here("Data", "site5", "5_4_points.shp"))
pts_5_5 <- vect(here("Data", "site5", "5_5_points.shp"))
```


Import site 2 rasters for prediction
```{r}
rast_5_2 <- rast(here("Data", "site5", "5_2_rast.tif"))
rast_5_3 <- rast(here("Data", "site5", "5_3_rast.tif"))
rast_5_4 <- rast(here("Data", "site5", "5_4_rast.tif"))
rast_5_5 <- rast(here("Data", "site5", "5_5_rast.tif"))
```

Rename raster layers
```{r}
subsite_ids <- 2:5
for (subsite_id in subsite_ids) {
  raster <- get(paste0("rast_5_", subsite_id))
  names(raster) <- c("red", "green", "blue")
  assign(paste0("rast_5_", subsite_id), raster)
  print(paste("Renamed layers for site 5_", subsite_id))
}
```
Create vector to store accuracy results
```{r}
site5_acc <- numeric(length(subsite_ids))
names(site5_acc) <- paste0("site_5_", subsite_ids)
```

Create a loop that classifies each raster site, extracts the classnames at certain points, and compares them to the known classnames of those points to obtain classification accuracy. 
```{r}
for (i in seq_along(subsite_ids)) {
  subsite_id <- subsite_ids[i]
  raster_to_classify_5 <- get(paste0("rast_5_", subsite_id))
  pts_shapefile_5 <- get(paste0("pts_5_", subsite_id))
  classified_raster_5 <- predict_spatial(raster_to_classify_5, trainSet1_lrn)
  classified_file_path_5 <- here("Data", "site5", paste0("5_", subsite_id, "_classified.tif"))
  writeRaster(classified_raster_5, classified_file_path_5, overwrite=T)
  extracted_vals_5 <- extract(classified_raster_5, pts_shapefile_5)
  raster_className <- extracted_vals_5[,2]
  shapefile_className <- pts_shapefile_5$className
  site5_acc_vect <- raster_className == shapefile_className
  class_acc_5 <- sum(site5_acc_vect) / length(site5_acc_vect)
  site5_acc[i] <- class_acc_5
  print(paste("Classification Accuracy for site5_", subsite_id, ":", class_acc_5))
}
print(site5_acc)
```

#Obtain classification accuracy of other subsites in site 6
Import site 2 points with known classes
```{r}
pts_6_2 <- vect(here("Data", "site6", "6_2_points.shp"))
pts_6_3 <- vect(here("Data", "site6", "6_3_points.shp"))
pts_6_4 <- vect(here("Data", "site6", "6_4_points.shp"))
pts_6_5 <- vect(here("Data", "site6", "6_5_points.shp"))
```


Import site 2 rasters for prediction
```{r}
rast_6_2 <- rast(here("Data", "site6", "6_2_rast.tif"))
rast_6_3 <- rast(here("Data", "site6", "6_3_rast.tif"))
rast_6_4 <- rast(here("Data", "site6", "6_4_rast.tif"))
rast_6_5 <- rast(here("Data", "site6", "6_5_rast.tif"))
```

Rename raster layers
```{r}
subsite_ids <- 2:5
for (subsite_id in subsite_ids) {
  raster <- get(paste0("rast_6_", subsite_id))
  names(raster) <- c("red", "green", "blue")
  assign(paste0("rast_6_", subsite_id), raster)
  print(paste("Renamed layers for site 6_", subsite_id))
}
```
Create vector to store accuracy results
```{r}
site6_acc <- numeric(length(subsite_ids))
names(site6_acc) <- paste0("site_6_", subsite_ids)
```

Create a loop that classifies each raster site, extracts the classnames at certain points, and compares them to the known classnames of those points to obtain classification accuracy. 
```{r}
for (i in seq_along(subsite_ids)) {
  subsite_id <- subsite_ids[i]
  raster_to_classify_6 <- get(paste0("rast_6_", subsite_id))
  pts_shapefile_6 <- get(paste0("pts_6_", subsite_id))
  classified_raster_6 <- predict_spatial(raster_to_classify_6, trainSet1_lrn)
  classified_file_path_6 <- here("Data", "site6", paste0("6_", subsite_id, "_classified.tif"))
  writeRaster(classified_raster_6, classified_file_path_6, overwrite=T)
  extracted_vals_6 <- extract(classified_raster_6, pts_shapefile_6)
  raster_className <- extracted_vals_6[,2]
  shapefile_className <- pts_shapefile_6$className
  site6_acc_vect <- raster_className == shapefile_className
  class_acc_6 <- sum(site6_acc_vect) / length(site6_acc_vect)
  site6_acc[i] <- class_acc_6
  print(paste("Classification Accuracy for site6_", subsite_id, ":", class_acc_6))
}
print(site6_acc)
```

#Obtain classification accuracy of other subsites in site 7
Import site 2 points with known classes
```{r}
pts_7_2 <- vect(here("Data", "site7", "7_2_points.shp"))
pts_7_3 <- vect(here("Data", "site7", "7_3_points.shp"))
pts_7_4 <- vect(here("Data", "site7", "7_4_points.shp"))
pts_7_5 <- vect(here("Data", "site7", "7_5_points.shp"))
```


Import site 2 rasters for prediction
```{r}
rast_7_2 <- rast(here("Data", "site7", "7_2_rast.tif"))
rast_7_3 <- rast(here("Data", "site7", "7_3_rast.tif"))
rast_7_4 <- rast(here("Data", "site7", "7_4_rast.tif"))
rast_7_5 <- rast(here("Data", "site7", "7_5_rast.tif"))
```

Rename raster layers
```{r}
subsite_ids <- 2:5
for (subsite_id in subsite_ids) {
  raster <- get(paste0("rast_7_", subsite_id))
  names(raster) <- c("red", "green", "blue")
  assign(paste0("rast_7_", subsite_id), raster)
  print(paste("Renamed layers for site 7_", subsite_id))
}
```
Create vector to store accuracy results
```{r}
site7_acc <- numeric(length(subsite_ids))
names(site7_acc) <- paste0("site_7_", subsite_ids)
```

Create a loop that classifies each raster site, extracts the classnames at certain points, and compares them to the known classnames of those points to obtain classification accuracy. 
```{r}
for (i in seq_along(subsite_ids)) {
  subsite_id <- subsite_ids[i]
  raster_to_classify_7 <- get(paste0("rast_7_", subsite_id))
  pts_shapefile_7 <- get(paste0("pts_7_", subsite_id))
  classified_raster_7 <- predict_spatial(raster_to_classify_7, trainSet1_lrn)
  classified_file_path_7 <- here("Data", "site7", paste0("7_", subsite_id, "_classified.tif"))
  writeRaster(classified_raster_7, classified_file_path_7, overwrite=T)
  extracted_vals_7 <- extract(classified_raster_7, pts_shapefile_7)
  raster_className <- extracted_vals_7[,2]
  shapefile_className <- pts_shapefile_7$className
  site7_acc_vect <- raster_className == shapefile_className
  class_acc_7 <- sum(site7_acc_vect) / length(site7_acc_vect)
  site7_acc[i] <- class_acc_7
  print(paste("Classification Accuracy for site7_", subsite_id, ":", class_acc_3))
}
print(site7_acc)
```
Print all accuracy results
```{r}
print(site1_acc)
print(site2_acc)
print(site3_acc)
print(site4_acc)
print(site5_acc)
print(site6_acc)
print(site7_acc)
```





